# Parameterized COVID-19 report 

Parameterized reporting is a technique that allows you to generate multiple reports simultaneously. By using parameterized reporting, you can follow the same process to make 3,000 reports as you would to make one report. The technique also makes your work more accurate, as it avoids copy-and-paste errors.

In this chapter we're making a parameterized report of COVID-19 data obtained from the European Center for Disease Control (ECDC). 


## Importeren en inspecteren {-}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
# importeren data
covid_data <- read_csv("~/Rachelle/data.csv")

library(reactable)
reactable(covid_data, defaultPageSize =10, compact = TRUE)
```

## Line chart the Netherlands, Belgium and Germany {-}
Using the ECDC COVID-19 data, there is a visual representation of COVID-19 cases and deaths in selected countries and years. The graphs display the number of cases and deaths per date. By utilizing parameterization in the R code with params$, this report can be easily tailored to specific interests without needing to modify the underlying code.
```{r, echo=F, message=F, fig.cap= "Line chart of the COVID-19 data from the Netherlands, Belgium and Germany, with the date in years and month on t x-axis, and the amount of cases on the y-axis. de datum in jaar en maand op de x-as staat, de aantal cases op de y-as."}

library(ggplot2)

# parameters
param <- list(
  selecteer_land= c('the Netherlands', 'Germany', 'Belgium'),
  selecteer_jaar= 2020
)

# converteer dateRep naar het Date-type
covid_data$dateRep <- as.Date(covid_data$dateRep, format='%d/%m/%Y')

# filter de data voor geselecteerde landen
selecteer_data <- subset(covid_data, countriesAndTerritories %in% param$selecteer_land & format(dateRep, "%Y") == param$selecteer_jaar)

# aggregeer de data per datum en land
agg_data <- aggregate(cbind(cases, deaths) ~ dateRep + countriesAndTerritories, data=selecteer_data, sum)

# grafiek cases
cases_plot <- ggplot(agg_data, aes(x=dateRep, y=cases, color=countriesAndTerritories)) +
  geom_line() +
  labs(title=paste('COVID-19 cases in selected countries in time', param$selecteer_jaar), x='Date', y=' amount of cases') +
  theme_minimal()

# grafiek deaths
deaths_plot <- ggplot(agg_data, aes(x=dateRep, y=deaths, color=countriesAndTerritories)) +
  geom_line() +
  labs(title=paste('COVID-19 deaths in selected countries in time', param$selecteer_jaar), x='Date', y='amount of cases') +
  theme_minimal()

# plot grafieken
print(cases_plot)
print(deaths_plot)

```

## Line chart Romania, Liechtenstein and Italy {-}
To verify that the parameterization is working correctly, you can adjust the values in the parameter list. In this example, the countries Romania, Norway, and Italy are used, along with the year 2022
```{r, echo=F, message=F, fig.cap= "Line chart of COVID-19 data, with the date in year and month on the x-axis, the number of cases on the y-axis. Different colors represent the selected countries Romania, Liechtenstein, and Italy."}
library(ggplot2)

# parameters
param <- list(
  selecteer_land= c('Romania', 'Liechtenstein', 'Italy'),
  selecteer_jaar= 2022
)

# converteer dateRep naar het Date-type
covid_data$dateRep <- as.Date(covid_data$dateRep, format='%d/%m/%Y')

# filter de data voor geselecteerde landen
selecteer_data <- subset(covid_data, countriesAndTerritories %in% param$selecteer_land & format(dateRep, "%Y") == param$selecteer_jaar)

# aggregeer de data per datum en land
agg_data <- aggregate(cbind(cases, deaths) ~ dateRep + countriesAndTerritories, data=selecteer_data, sum)

# grafiek cases
cases_plot <- ggplot(agg_data, aes(x=dateRep, y=cases, color=countriesAndTerritories)) +
  geom_line() +
  labs(title=paste('COVID-19 cases in slected countries in time', param$selecteer_jaar), x='Date', y='Amount of cases') +
  theme_minimal()

# grafiek deaths
deaths_plot <- ggplot(agg_data, aes(x=dateRep, y=deaths, color=countriesAndTerritories)) +
  geom_line() +
  labs(title=paste('COVID-19 cases in slected countries in time', param$selecteer_jaar), x='Date', y='Amount of cases') +
  theme_minimal()

# plot grafieken
print(cases_plot)
print(deaths_plot)

```

