# <span style="color:#E656AE;">Parameterized COVID-19 report</span> 

Parameterized reporting is a technique that allows you to generate multiple reports simultaneously. By using parameterized reporting, you can follow the same process to make 3,000 reports as you would to make one report. The technique also makes your work more accurate, as it avoids copy-and-paste errors.

In this chapter we're making a parameterized report of COVID-19 data obtained from the European Center for Disease Control (ECDC). 


## Importing and inspecting  {-}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
# importeren data
covid_data <- read_csv("~/Rachelle/data.csv")

library(reactable)
reactable(covid_data, defaultPageSize =10, compact = TRUE)
```

## Line chart the Netherlands, Belgium and Germany {-}
Using the ECDC COVID-19 data, there is a visual representation of COVID-19 cases and deaths in selected countries and years. The graphs display the number of cases and deaths per date. By utilizing parameterization in the R code with `params$`, this report can be easily tailored to specific interests without needing to modify the underlying code.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# parameters
params <- list(
  selecteer_land = c("Netherlands", "Germany", "Belgium"),
  selecteer_jaar = 2020
)

# Datum omzetten
selecteer_data <- covid_data %>%
  mutate(dateRep = as.Date(dateRep, format = "%d/%m/%Y")) %>%
  filter(
    countriesAndTerritories %in% params$selecteer_land,
    format(dateRep, "%Y") == as.character(params$selecteer_jaar)
  )


plot_data <- selecteer_data %>%
  select(dateRep, countriesAndTerritories, cases, deaths) %>%
  pivot_longer(cols = c(cases, deaths), names_to = "type", values_to = "aantal")

# Plot 
ggplot(plot_data, aes(x = dateRep, y = aantal, color = type)) +
  geom_line() +
  facet_wrap(~countriesAndTerritories) +
  labs(
    title = paste("COVID-19 cases and deaths in", params$selecteer_jaar),
    x = "Date", y = "Count", color = "Type"
  ) +
  theme_minimal() + scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Line chart Romania, Liechtenstein and Italy {-}
To verify that the parameterization is working correctly, you can adjust the values in the parameter list. In this example, the countries Romania, Norway, and Italy are used, along with the year 2022

```{r message=FALSE, warning=FALSE}
# Laad packages
library(tidyverse)

# parameters
params <- list(
  selecteer_land = c('Romania', 'Liechtenstein', 'Italy'),
  selecteer_jaar = 2022
)

plot_data <- covid_data %>%
  mutate(dateRep = as.Date(dateRep, format = "%d/%m/%Y")) %>%
  filter(
    countriesAndTerritories %in% params$selecteer_land,
    format(dateRep, "%Y") == as.character(params$selecteer_jaar)
  ) %>%
  select(dateRep, countriesAndTerritories, cases, deaths) %>%
  pivot_longer(cols = c(cases, deaths), names_to = "type", values_to = "aantal")

# Grafiek maken met cases en deaths 
ggplot(plot_data, aes(x = dateRep, y = aantal, color = countriesAndTerritories)) +
  geom_line() +
  facet_wrap(~type, scales = "free_y") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  labs(
    title = paste("COVID-19 cases and deaths in", params$selecteer_jaar),
    x = "Date", y = "Count", color = "Country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

